### 2015-08-09 manuscript/abstract.tex
Paper files.

\begin{abstract}
This paper proposes an application that collects sensor data from smartwatches and smartphones in order to predict blood alcohol content (BAC) or drunkenness in real-time, discreetly, and non-invasively through a machine learning approach. This system could prevent drunk driving or other dangers related to the consumption of alcohol by giving users a way to measure personal BAC without the use of breathalyzers or other typically undesired methods. Using data collected by a custom application from many volunteers, we trained a machine learning model that will work with an Android application to predict the user’s BAC.
\end{abstract}

\begin{IEEEkeywords}
	\lowercase{Sensor systems and applications, Temperature sensors, Machine learning, Prediction methods, Prediction algorithms, Internet of Things, Middleware, Wireless sensor networks}
\end{IEEEkeywords}