### 2015-08-16 manuscript/content/abstract.tex
Switch format and other changes

\begin{abstract}
This paper proposes an application that collects sensor data from smartwatches and smartphones in order to predict blood alcohol content (BAC) or drunkenness in real-time, discreetly, and non-invasively through a machine learning approach. This system could prevent drunk driving or other dangers related to the consumption of alcohol by giving users a way to measure personal BAC without the use of intrusive breathalyzers or guess work.  Using data collected by a custom application from many volunteers, we trained a machine learning model that will work with an Android application to predict the user's BAC in real-time.
\end{abstract}

%\begin{IEEEkeywords}
	%\lowercase{Sensor systems and applications, Temperature sensors, Machine learning, Prediction methods, Prediction algorithms, Internet of Things, Middleware, Wireless sensor networks}
%\end{IEEEkeywords}