### 2015-08-24 manuscript/content/system.tex

\section{System Architecture}

% Present the overall system architecture here.

\subsection{Android-based Sensor Gateway}

% Describe the sensor gateway here.